Multi-redshift limits on the 21cm power spectrum from PAPER

### 2014-11-10 ms.tex
revised 4.3 to be more clear about what we think the excess is

A grid configuration is sufficiently different from a traditional interferometer antenna layout that 
there are many tradeoffs to consider. In this configuration foregrounds cannot be imaged, and therefor can only be filtered or blindly subtracted.  Experimentally, the loss of foreground imaging is offset by an enhanced experimental precision. The grid configuration makes repeated measurements of the \emph{same} modes with \emph{independent} antennas, thus the per-antenna gains are a tractable linear algebra problem that does not depend on a sky model \citep{Liu:2010p10391}. This same redundancy also enables the identification of systematically deviant samples, allows us to avoid noise bias, and provides another axis along which to bootstrap.  These are believed to provide the greatest handle on known systematics.  This solution has a small number of free parameters which must be calibrated to a model of the sky to obtain a power spectrum calibrated to a standard flux scale.  Flux and overall bandpass calibration is the area most impacted by the small number of sampled spatial modes.  In practice, the use of synthesis rotation allows for adequate coverage of the brightest source in the sky which we use as our flux standard (see section \ref{sec:calibration}.

@@

\subsection{The Excess Power}
\label{sec:excess power}
As these measurements have a maximum possible sensitivity level which is still two orders of magnitude away from theoretical reionization levels we consider the likelihood that these detections are evidence of reionization to be small. Far more likely is that we are limited by a residual foreground or systematic arising from approximations in the data analysis.  The statistical significance of the residual --2$\sigma$ after averaging over the entire season-- is low enough that it is not possible to break it down along likely axes like sidereal time or time of day to probe its origin.  Such studies must await higher sensitivity measurements with more antennas.  However, there are several additional features that seem to suggest an excess not caused by high redshift HI.

The spectral signature of the measured excess is far more similar to a foreground residual then to theoretical HI models.
The predicted signature of the HI signal during reionization is one of a relatively short burst of power at small $k$ values as the ionized bubbles are briefly at their largest, before the signal from the ever increasing ionized medium drops sharply \citep{Pritchard:2008p8123}. In contrast, this observed residual is relatively constant with redshift while rising at both ends of the band, mimicking the RMS noise curve in Figure \ref{fig:noise}.  This suggests that the accuracy of the foreground filter decreases towards the edges of the band.  Recall that the wide-band foreground model we have subtracted was built by weighting the entire band by a Blackman-Harris window function which provides a much higher isolation of foregrounds inside the horizon. The tradeoff is that data in the outer parts of the spectrum are heavily down weighted.  For, in the subtracted foreground model, data at 126MHz are down-weighted $\sim$75\% compared to 150Mhz.  This amounts to making an assumption about the spectral evolution of the data, namely that 126MHz can be mostly modeled by extrapolation from 150MHz.  The precision of this subtraction can be judged by examining residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.  The relative depth of the redshift 10.5 P(k) foreground residual is substantially shallower then at 7.55 and is generally correlated with the amount of residual found at higher $k$s.

@@

\section{Acknowledgements}
The PAPER project is supported through the NSF-AST program (award \#1129258). Computing resources were provided by a grant from Mt. Cuba Astronomical Foundation.  J.C.P. is supported by an NSF Astronomy and Astrophysics Fellowship under award AST \#1302774.   We thank Paul Sutter and Jonathan Pritchard for helpful discussions.

>>>>

A grid configuration is sufficiently different from a traditional interferometer antenna layout that 
there are many tradeoffs to consider. In this configuration foregrounds cannot be imaged, and therefor can only be filtered or subtracted by forward modeling.  Experimentally, the loss of foreground imaging is offset by an enhanced experimental precision. The grid configuration makes repeated measurements of the \emph{same} modes with \emph{independent} antennas, thus the per-antenna gains are a tractable linear algebra problem that does not depend on a sky model \citep{Liu:2010p10391}. This same redundancy also enables the identification of systematically deviant samples, allows us to avoid noise bias, and provides another axis along which to bootstrap.  These are believed to provide the greatest handle on known systematics.  This solution has a small number of free parameters which must be calibrated to a model of the sky to obtain a power spectrum calibrated to a standard flux scale.  Flux and overall bandpass calibration is the area most impacted by the small number of sampled spatial modes.  In practice, the use of synthesis rotation allows for adequate coverage of the brightest source in the sky which we use as our flux standard (see section \ref{sec:calibration}.

@@

\subsection{The Excess Power}
\label{sec:excess power}
%As these measurements have a maximum possible sensitivity level which is still two orders of magnitude away from theoretical reionization levels, we consider the likelihood that these detections are evidence of reionization to be small. Far more likely, is that we are limited by a residual foreground or systematic arising from approximations in the data analysis.
  The statistical significance of the residual --$<2\sigma$ for most points-- is low enough that it is not possible to investigate possible sources of systematic bias --e.g. dependence on sidereal time, time of day, or other dimensions.  Such studies must await higher sensitivity measurements with more antennas. Without data sensitive enough to dissect the residuals we cannot identify their origin, but we can make a couple observations.

The redshift 7.68 bin reproduces the excess seen in P14 where it was observed that a complete lack of X-Ray heating from early stars would be required to produce a signal two orders of magnitude above expected theoretical levels.  Extending this analysis to the spectral dimension we find more significant deviation from theoretical behavior.  The theoretical signature of the HI signal during reionization is one of a relatively short burst of power at small $k$ values, as the ionized bubbles are briefly at their largest, before the signal from the ever increasing ionized medium drops sharply \citep{Pritchard:2008p8123}.  The observed residual is relatively constant with redshift, rising at both ends of the band, mimicking the RMS noise curve in Figure \ref{fig:noise}.  This suggests that the accuracy of the foreground filter decreases towards the edges of the band.  Recall that the wide-band foreground model we have subtracted was built by weighting the entire band by a Blackman-Harris window function which provides a much higher isolation of foregrounds inside the horizon. The tradeoff is that data in the outer parts of the spectrum are heavily down weighted.  For example, in the subtracted foreground model, data at 126\,MHz are down-weighted $\sim$75\% compared to 150\,Mhz.  This amounts to making an assumption about the spectral evolution of the data, namely that 126\,MHz can be mostly modeled by extrapolation from 150\,MHz.  The precision of this subtraction can be judged by examining residuals inside the horizon in the $P(k)$ power spectra plotted in the upper part of Figure \ref{fig:pspecs}\footnote{Note that this notch is not present in Figure \ref{fig:P14compare} because the foregrounds have been left in for comparison with P14.}.  The relative depth of the redshift 10.5 $P(k)$ foreground residual is substantially shallower then at 7.55 and is generally correlated with the amount of residual found at higher $k$s.

@@

\section{Acknowledgements}
The PAPER project is supported through the NSF-AST program (award \#1129258). Computing resources were provided by a grant from Mt. Cuba Astronomical Foundation.  D.C.J. is supported by an NSF Astronomy and Astrophysics Fellowship under award AST \# 1401708 while J.C.P. is supported by an NSF Astronomy and Astrophysics Fellowship under award AST \#1302774.   We thank Paul Sutter and Jonathan Pritchard for helpful discussions.
